---
title: "Practical Machine Learning Course Project"
author: "Thomas Tong"
date: "Saturday, December 20, 2014"
output: html_document
---

## Load Required Packages

Load RCurl and caret libraries. Other depending libraries will be loaded when they are needed.

```{r}
library(RCurl)
library(caret)
```

## Load Data and Preprocessing

Load training data and testing data from HTTPS source. Treat characters values "#DIV/0!", "NA" as NA.

```{r}
trainUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
trainData <- read.csv(text=getURL(trainUrl, ssl.verifypeer=0L, followlocation=1L), 
                      na.strings=c("#DIV/0!", "NA"))

testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
testData <- read.csv(text=getURL(testUrl, ssl.verifypeer=0L, followlocation=1L), 
                     na.strings=c("#DIV/0!", "NA"))
```

The first seven columns are not related to our analysis.

```{r}
names(testData)[c(1:7)]
```

Drop the first seven columns as a part of our preprocessing.

```{r}
trainData <- trainData[,-(1:7)]
testData <- testData[,-(1:7)]
```

These columns are all NA in values.

```{r}
columnWithoutData <- apply(trainData,2,function(x) { sum(is.na(x)) })
names(testData)[which(columnWithoutData == 0)]
```

Remove columns with no meaningful data in order to avoid them affecting training.

```{r}
trainData <- trainData[,which(columnWithoutData == 0)]
testData <- testData[,which(columnWithoutData == 0)]
```

## Perform Data Partition

Set the seed for reproducible results. 

```{r}
set.seed(454556)
```

Seperate the training data into 70% training partition and 30% cross validation partition.

```{r}
dataPartition <- createDataPartition(y = trainData$classe, p=0.7, list=F)
trainPart <- trainData[dataPartition,]
crossValidPart <- trainData[-dataPartition,]
```

## Training

Perform random forest training on the training data partition.

```{r}
model <- train(classe ~., data = trainPart, method="rf")
model
```

Use the model to perform prediction on the cross validation data partition and use confusion matrix function to check the accuracy and error rate.

```{r}
crossValid <- predict(model, crossValidPart)
confusionMatrix(crossValid, crossValidPart$classe)
```

As we can see, the error rate is very low. It seems to us that the random forest training has done a pretty nice job in this analysis.

Then we use the model to perform prediction on test dataset.

```{r}
predition <- predict(model, testData)
predition
```